Deploying and Sharing QSAR Models using Predictive Modeling Markup Language

### 2014-12-02 paper.tex
Categorization of predictive solutions from the end user perspective

>>>>

\subsection{PMML export}

Completeness of QSAR predictive solutions:
\begin{enumerate}
  \item A model, without any extras. The default option for many popular 
  statistics applications (eg. the result of a "Save Model As ..."
  command). Model input and output fields are underspecified, typically
  limited only to the symbolic name of the field.
  Very hard for end users to work with. If the training dataset is available,
  could be even worthwhile to re-train the model (possibly using more PMML
  capable application) and export the result with more (meta-)information so that
  it would qualify for the second or thrid level (see below). 
  \item A model or a model ensemble, with basic data schema, basic data pre- 
  and post-processing. Selected statistics applications such as R and KNIME
  provide facilities for specifying data pre- and post-processing functions
  using PMML transformations (\url{http://www.dmg.org/v4-2-1/Transformations.html}).
  Today, KNIME can be considered to be the most powerful statistics 
  application for generating custom PMML documents. For example, KNIME version
  2.9 and newer provide graphical editor for assembling workflows that produce
  model ensembles.
  Can be deployed by at least moderately tech-savvy end users. The biggest
  challenge is connecting the solution to the actual data source. For example,
  if the model was trained in R using a data.frame of CDK descriptors, then
  the end-user must be able to calculate CDK descriptor values for the
  "incoming" chemical compounds.
  \item An "interaction" of models, with advanced data schema, data pre- 
  and post-processing. The PMML document is produced programmatically by
  custom software application. Could use a "template" approach, where the base
  of a PMML document is hand-crafted, and model placeholder elements are
  subsituted with fragment PMML documents that are produced by ordinary
  statistics applications. The practical experience of one of the authors (VR)
  suggests that the template approach is easy to set up and maintain even
  if the overall familiarity with PMML in the organization is not that high.
  Another customization object is reshaping the data schema of a predictive
  solution to expected end-user needs. Moving re-defining the input fields
  so that it becomes possible of inputting a single SMILES string instead of
  50+ CDK descriptor values.
\end{enumerate}
